---
title: "Explorative DataViz Short Paper"
author: "Ji, Ye"
date: "10/15/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Assignment Requirement
This assignment builds off the last assignment, resulting in a single document either in PDF or HTML form exported from an R markdown document. You must use the data exclusively from your portfolio proposal! Further, this document should have:

* Five data visualizations of most distinct graph types, made with ggplot2; This can include the three visualizations from last week, though you can replace them if you want. Note that graph requirements always include:

    + Correct usage of all visual encodings;

    + Appropriate data sourcing;

    + Proper labeling of ALL visual encodings, as well as an appropriate title and subtitle;

    + The code you used to generate each graphic, right above the graph (R Markdown should make this easy). 

* A written narrative for each graph, from three to five sentences, describing the context of the graph and how it informs about your policy topic. You can write this as a narrative.

Note that, over time, you are expected to tackle progressively more varied and ambitious graph types. At this stage in the course, the standard is still not very high. That said, you should be able to clearly articulate the value of the graphs you created so far - meaning they need to make sense and have some value-add. 

Please submit the URL of the file on a Git repository (it can be one git repo you keep adding to, or a stand alone repository).

```{r, echo=TRUE, include=FALSE, results="hide"}
setwd("/Users/apple/Github/data_viz_works/")

install.packages("easypackages", repos = "http://cran.us.r-project.org")
library(easypackages)
packages("anytime","tidyverse","rmarkdown","readr", "haven", "dplyr", "stringr", "ggplot2", "xts", "igraph")

# Import data
coinbase = read_csv("./data/coinbaseUSD.csv")
colnames(coinbase) = c("unix_timestamp", "price", "volume_btc")
                         
# Convert unix timestamp to human-readable date & time
coinbase <- mutate(coinbase, unix_timestamp = anytime(unix_timestamp))
coinbase <- mutate(coinbase, volume_usd = volume_btc * price)

# sample down to 1 percent
coinbase_lim <- sample_frac(coinbase, size = 0.01)

# aggregate trading volume data into month
ts <- xts(coinbase_lim$volume_usd, as.Date(coinbase_lim$unix_timestamp, "%Y-%m-%d"))
ts_d = apply.daily(ts, FUN=sum)
ts_m = apply.monthly(ts_d, FUN=sum)
df_m <- data.frame(date=index(ts_m), coredata(ts_m))
```
### Introduction

There are several advantages to dedicate a data visualization portfolio and 10-week effort to digital currency:

    + Massive and Robust data source: Blockchain is a public transaction log
    that exists as a distributed database, validated by its powerful algorithm.
    
    + Fitting data types: Blockchain data includes time and IP addresses, and 
    can be updated frequently, which means it is useful for both the static and
    dynamic data visualizations.
    
    + Well-supported: I can save the entire blockchain file locally,use 
    community-maintained tools to parse it, and keep the data up-to-date.
    
    + Policy-setting: Digital currencies are being used as store of value to 
    fight inflation in turbulent economies such as Zimbabwe and Latin American
    countries. Central Banks are issuing experimental policies in response to 
    the dessemination of digital currency and blockchain technology.
    
While I'm waiting for my local blockchain file (150 GB and growing) to finish download, I will explore some secondary data sources published by major Bitcoin exchanges and platforms to better understand the data.


\pagebreak

### Graph one - line chart
```{r, fig.width = 16, fig.height = 9, result="hide"}
line <- ggplot(data=coinbase_lim) +
  geom_line(size=.25,
            aes(unix_timestamp, price),
            color="#325a8c") +
  geom_vline(xintercept = as.POSIXct("2017-03-25"),
             colour="#ff7575",
             size=.75,
             alpha=.75) +
  geom_smooth(aes(unix_timestamp,price),
              span=2,
              linetype="longdash",
              color="#4c88d3",
              alpha=.5) +
  scale_x_datetime(name = "",
                   date_breaks = "3 month") +
  scale_y_continuous(name ="Bitcoin Hourly Exchange Rate in USD",
                     breaks = c(0, 500, 1000, 1500, 2000, 2500, 3000)) +
  ggtitle("Bitcoin Daily Price",
          subtitle = "experienced the greatest surge in history this year") +
  labs(caption = "Data source: Coinbase API") +
  annotate(geom="text",
           x=as.POSIXct("2017-05-30"),
           y=675,
           label="2017-03-25 Surge starts",
           colour="#ff7575",
           fontface="bold",
           alpha=.85) +
  theme(panel.grid.major.y = element_line( size=.1, color="#666666"),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size=18,
                                  family = "Helvetica",
                                  colour = "#3a3a3a",
                                  face = "bold"),
        plot.subtitle = element_text(size=12, family = "mono", colour = "#666666"),
        axis.title.y.right = element_text(color="#85bb65"),
        axis.text.y.right = element_text(color="#85bb65"),
        axis.text.y = element_text(color="#325a8c"),
        axis.title.y = element_text(color="#325a8c"),
        axis.ticks.y = element_line( size=.25, color="#666666"),
        axis.text.x = element_text(color="#7f7f7f"),
        legend.position = "none",
        plot.caption = element_text(size=11, family = "mono", colour = "#666666"),
        plot.margin = unit(c(2,2,2,2), "cm"))

line
```

# Narrative:
The price of bitcoin can increase or decrease drastically over a
short period of time. This graph mainly serves as a priliminary 
clue to pinpoint certain time points where a new policy may have
impacted the price of Bitcoin, or vice versa. And potentially, 
this graph could be used to find global price trend correlations.


\pagebreak

### Graph two - histogram
```{r, echo=FALSE, include=FALSE, result="hide"}
nodes <- read_csv("./data/nodes_by_countries.csv")
```

```{r, fig.width = 12, fig.height = 18}
hist <- ggplot(data=nodes, aes(reorder(factor(country), node), node)) +
  geom_col(width=0.9, position=position_dodge(width=5), fill = "#999999") +
  geom_text(aes(label=node),
            color="#666666",
            position=position_dodge(width=0.9),
            vjust=0.35,
            hjust=-.25) +
  ggtitle("Bitcoin Nodes Distribution by Countries",
          subtitle = "Reachable nodes as of Oct 15, 2017") +
  ylab("") +
  xlab("") +
  labs(caption = "Data source: BitNodes.21.co") +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size=18,
                                  family = "Helvetica",
                                  colour = "#3a3a3a",
                                  face = "bold"),
        plot.subtitle = element_text(size=12, family = "mono", colour = "#666666"),
        axis.text.y = element_text(color="#325a8c"),
        axis.title.y = element_text(color="#325a8c"),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none",
        plot.caption = element_text(size=11, family = "mono", colour = "#666666"),
        plot.margin = unit(c(2,2,2,2), "cm")) +
  scale_x_discrete(expand=c(0,0)) +
  coord_flip()

hist
```

# Narrative:
Countries that have more reachable nodes means more people in
the countries are involved in Bitcoin mining, which indicates
that these countries are more likely to have policies,
or the lack of, to support the development of cryptocurrency.
For an explorative analysis, this graph is helpful for narrowing
down the origin of some major policies that have shaped the 
development of Bitcoin. If we consider this graph as a ranking
of countries' involvement in digital currency. It is interesting
to note that, this ranking is not necessarily ordered in exact 
accordance with the ranking of GDP, population, technological
advancement, or even perceived internet presence.

\pagebreak

## Graph Three - stacked area graph
```{r, echo=FALSE, include=FALSE, result="hide"}
ts <- xts(coinbase_lim$volume_usd, as.Date(coinbase_lim$unix_timestamp, "%Y-%m-%d"))
ts_d = apply.daily(ts, FUN=sum)
#ts_m = apply.monthly(ts_d, FUN=sum)
df_d <- data.frame(date=index(ts_d), coredata(ts_d))
colnames(df_d) <- c("date", "volume_usd")

ts_btc <- xts(coinbase_lim$volume_btc, as.Date(coinbase_lim$unix_timestamp, "%Y-%m-%d"))
ts_d_btc = apply.daily(ts_btc, FUN=sum)
#ts_m_btc = apply.monthly(ts_d_btc, FUN=sum)
df_d_btc <- data.frame(date=index(ts_d_btc), coredata(ts_d_btc))
colnames(df_d_btc) <- c("date", "volume_btc")

df <- left_join(df_d, df_d_btc)

df_btc <- df[, c("date", "volume_btc")]
df_btc$unit <- "Bitcoin"
colnames(df_btc) <- c("date", "volume", "unit")

df_usd <- df[, c("date", "volume_usd")]
df_usd <- mutate(df_usd, volume_usd = volume_usd/1000)
df_usd$unit <- "Thousand USD"
colnames(df_usd) <- c("date", "volume", "unit")

agg <- rbind(df_btc, df_usd)
```


```{r, fig.width = 16, fig.height = 12}
# note: color #85bb65 is nicknamed as "dollar bill green".
#       color #FF9900 is the orange from bitcoin logo.
area <- ggplot(data=agg, aes(x = as.POSIXct(date), y = volume, fill = as.factor(unit))) +
  geom_area(position = "stack") +
  scale_x_datetime(date_breaks = "3 months", expand=c(0,0)) +
  scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 2000, 3000)) + 
  scale_fill_manual(values = alpha(c("#FF9900", "#85bb65")), 0.2) +
  ggtitle("Bitcoin Daily Trading Volume",
          subtitle = "in number of bitcoin & Thousands of USD") +
  labs(caption = "Data source: Coinbase API",
       x = "",
       y = "Volume") +
  theme(panel.grid.major.y =element_line( size=.1, color="#999999"),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size=18,
                                  family = "Helvetica",
                                  colour = "#3a3a3a",
                                  face = "bold"),
        plot.subtitle = element_text(size=12, family = "mono", colour = "#666666"),
        axis.text.y = element_text(color="#666666"),
        axis.title.y = element_text(color="#666666"),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(size=10, family = "mono", colour = "#666666"),
        legend.title = element_blank(),
        legend.position = c(0.1, 0.9),
        plot.caption = element_text(size=11, family = "mono", colour = "#666666"),
        plot.margin = unit(c(2,2,2,2), "cm"))

area
```

# Narrative:
In this graph, we can see that bitcoin's daily trading volume is extremely
volatile, perhaps even more so than its price. This level of volatility can
be a sign for market manipulation, especially given Bitcoin's young economy,
novel nature, and lack of regulation. Together with the price trend, this 
graph can help identify major policy changes that either encourage (spike)
or discourage (trough) speculation.

\pagebreak

## Graph Four - network of trust
```{r, echo=FALSE, include=FALSE, result="hide"}
soc <- read_csv("./data/soc-sign-bitcoinotc.csv", col_names = c("rater", "ratee", "score", "rate_timestamp"))
```

```{r, fig.width=14, fig.height=12}
# View in full screen for clarity.
heatmap <- ggplot(soc, aes(x=rater, y=ratee, fill=score)) + geom_tile() +
  scale_fill_distiller(type="div", palette = "RdYlBu") +
  ggtitle("The who-trusts-whom nework of Bitcoin trader on a platform called Bitcoin OTC",
          subtitle = "is sparse and unregulated") +
  labs(caption = "Data source: Bitcoin OTC trust network by Stanford University") +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size=18,
                                  family = "Helvetica",
                                  colour = "#3a3a3a",
                                  face = "bold"),
        plot.subtitle = element_text(size=12, family = "mono", colour = "#666666"),
        axis.text.y = element_text(color="#666666"),
        axis.title.y = element_text(color="#666666"),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(size=10, family = "mono", colour = "#666666"),
        legend.title = element_text(),
        legend.position = c(0, 0),
        plot.caption = element_text(size=11, family = "mono", colour = "#666666"),
        plot.margin = unit(c(2,2,2,2), "cm"))
heatmap
```

# Narrative:
While sparse and barely visible, this graph does provide some
insights about the Bitcoin market. First, it allows us to identify
some super raters and ratees, who are more likely to be the major
player capable of manipulating Bitcoin price. Second, although orange
(trustworthy) is more prevalent, there are quite a few blue spots 
on this heatmap, marking the users that are rated as fraud by others.
At a glimpse, fraud is frequent in digital currency market, which
underscores its unregulated nature.

\pagebreak

## Graph Five: long-term user network

# Narrative:
This is a graph that visualizes the transaction activities of several
thousand long-term Bitcoin users during 2009 - 2013. We can see that
the majority is inactive, with one user radiating assets to many others,
and a few others repeatedly send money to themselves. Not only can this
network graph serves as a piece of evidence to show that most people 
purcahse Bitcoin for the purpose of speculation, it also shows that, 
while the technological philosophy of Bitcoin is decentralization,
its ownership and commercial activity is very much the opposite.



### Graph Five: network
```{r, echo=FALSE, include=FALSE, result="hide"}
links <- read_tsv("./data/lt_graph_ELTE.txt", col_names = c("sender", "recipient", "transaction_timestamp"))
net <- graph_from_data_frame(d=links, directed = T)
```

```{r}
# ISSUE: cannot knit this chunk.

# set color
col.1 <- adjustcolor("#666666", alpha=0.4)
col.2 <- adjustcolor("#000000", alpha=0.4)
edge.pal <- colorRampPalette(c(col.1, col.2), alpha = TRUE)
edge.col <- edge.pal(100)

colrs_transparent <- adjustcolor("#FFFFFF", alpha=0)

#net_simplified <- simplify(net,remove.multiple = F,edge.attr.comb=list(Weight="sum"), remove.loops = T)
l <- layout_on_sphere(net)
plot(net,
     layout=l,
     edge.arrow.size=0.1,
     vertex.size=2,
     vertex.color="#FF9900",
     vertex.frame.color=colrs_transparent,
     vertex.label=NA,
     edge.color=edge.col,
     main="The majority of long-term Bitcoin users does not engage in transaction",
     sub="Data source: ELTE Bitcoin Project website and resources (2009 - 2013)",
     output.mode='inline'
     )
legend(x=-0.3, y=-1.1, c("long-term bitcoin user"), col="#FF9900", pch=21, pt.bg="#FF9900")
```
