---
title: "Theme & Design"
author: "Ji, Ye"
date: "10/20/2017"
output:
  html_document:
    fig_width: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE, include=FALSE}
# install packages.
install.packages("easypackages", repos = "http://cran.us.r-project.org")
library(easypackages)
packages("anytime","tidyverse","rmarkdown","readr", "haven", "dplyr", "stringr", "ggplot2", "xts", "igraph")

# load data
btc_price = read_csv("./data/BCHAIN-MKPRU.csv")
btc_price <- mutate(btc_price, Date = as.POSIXct(Date))

# set theme
theme_jiye <- theme(panel.grid.major.y =element_line( size=.1, color="#999999"),
                    panel.grid.major.x = element_blank(),
                    panel.background = element_blank(),
                    plot.title = element_text(size=24,
                                              family = "Helvetica",
                                              colour = "#3a3a3a",
                                              face = "bold"),
                    plot.subtitle = element_text(size=12,
                                                 family = "Avenir",
                                                 colour = "#666666"),
                    axis.title.y = element_text(colour="#325a8c"),
                    axis.title.x = element_blank(),
                    axis.text.y = element_text(colour="#325a8c"),
#                    axis.text.x = element_blank(),
                    axis.ticks.y = element_blank(),
#                    axis.ticks.x = element_blank(),
                    legend.position = "none",
                    plot.caption = element_text(size=8,
                                                family = "Avenir",
                                                colour = "#666666",
                                                hjust = 0
                                                ),
                    plot.margin = unit(c(1,1,1,1), "cm"))
```

### Introduction

There are several advantages to dedicate a data visualization portfolio and 10-week effort to digital currency:

* Massive and Robust data source: Blockchain is a public transaction log
  that exists as a distributed database, validated by its powerful algorithm.
    
* Fitting data types: Blockchain data includes time and IP addresses, and 
  can be updated frequently, which means it is useful for both the static and
  dynamic data visualizations.
    
* Well-supported: I can save the entire blockchain file locally,use 
  community-maintained tools to parse it, and keep the data up-to-date.
    
* Policy-setting: Digital currencies are being used as store of value to 
fight inflation in turbulent economies such as Zimbabwe and Latin American
countries. Central Banks are issuing experimental policies in response to 
the dessemination of digital currency and blockchain technology.
    
This paper will explore some data sources published by major Bitcoin exchanges and platforms as an market overview.


``` {r}
rectangle <- data.frame(xmin = as.POSIXct(c("2017-03-25")),
                        xmax = as.POSIXct(c("2017-10-21")),
                        ymin = -Inf,
                        ymax = Inf)

line <- ggplot(data=btc_price) +
  geom_line(size=.25,
            aes(Date, Value),
            color="#325a8c") +
  geom_vline(xintercept = as.POSIXct("2017-03-25"),
             colour="#ff7575",
             size=.5,
             linetype="dotted",
             alpha=.75) +
#  geom_smooth(aes(Date, Value),
#              size=0.5,
#              alpha=0.1,
#              span=0.5,
#              color="#325a8c",
#              linetype="twodash"
#              ) +
  scale_y_continuous(name ="Bitcoin Market Price in USD",
                     breaks = (seq(0,6000,1000))
                     ) +
  scale_x_datetime(date_breaks = "1 year",
                   labels = seq(2009,2018,1),
                   expand=c(0,0)) +
  ggtitle("Bitcoin Price",
          subtitle = "experienced the greatest surge in history this year") +
  labs(caption = "Source: Blockchain.com") +
  annotate(geom="text",
           x=as.POSIXct("2016-01-01"),
           y=2200,
           label="Surge started on Mar 25, 2017",
           colour="#ff7575",
           family="Avenir",
           fontface="bold",
           alpha=1) +
  theme_jiye

line + geom_rect(data = rectangle, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
                 fill = "#ff7575", alpha = 0.1)
```

<br/>

### Bitcoin price is volatile

The price of bitcoin can increase or decrease drastically over a
short period of time. This graph mainly serves as a priliminary 
evidence to pinpoint certain time points where a new policy or
event may have impacted the price of Bitcoin.

<br/>
