---
title: "Introduction to ggplot2: Three Charts"
author: "Ji, Ye"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Assignment Requirement
Explore your data with ggplot2, making and exporting to pdf three graphs of meaningfully distinct types. Submit those graphs to Canvas.

Use at least:

* five different aesthetics;

* five different non-aesthetic options;

* four different geoms;

* two different scales (meaning change scale option for at least two of the scales).

Include an narrative graph title, an explanatory subtitle, and explicit data sourcing. Make sure to include labels and scales for all visual encodings. 

```{r, echo=TRUE, include=FALSE, results="hide"}
#setwd("./GitHub/data_viz_works/DV-assignment2/")
install.packages("easypackages", repos = "http://cran.us.r-project.org")
library(easypackages)
packages("anytime","tidyverse","rmarkdown","readr", "haven", "dplyr", "stringr", "ggplot2", "gridExtra")
# Import data
coinbase = read_csv("coinbaseUSD.csv")
colnames(coinbase) = c("unix_timestamp", "price", "volume_btc")
                         
# Convert unix timestamp to human-readable date & time
coinbase <- mutate(coinbase, unix_timestamp = anytime(unix_timestamp))
coinbase <- mutate(coinbase, volume_usd = volume_btc * price)

# sample down to 1 percent
coinbase_lim <- sample_frac(coinbase, size = 0.01)

# aggregate trading volume data into month
packages("xts")
ts <- xts(coinbase_lim$volume_usd, as.Date(coinbase_lim$unix_timestamp, "%Y-%m-%d"))
ts_d = apply.daily(ts, FUN=sum)
ts_m = apply.monthly(ts_d, FUN=sum)
df_m <- data.frame(date=index(ts_m), coredata(ts_m))
```

### Chart one
```{r, fig.width = 16, fig.height = 9, result="hide"}
line <- ggplot(data=coinbase_lim) +
  geom_line(size=.25,
            aes(unix_timestamp, price),
            color="#325a8c") +
  geom_vline(xintercept = as.POSIXct("2017-03-25"),
             colour="#ff7575",
             size=.75,
             alpha=.75) +
  geom_smooth(aes(unix_timestamp,price),
              span=2,
              linetype="longdash",
              color="#4c88d3",
              alpha=.5) +
  scale_x_datetime(name = "",
                   date_breaks = "3 month") +
  scale_y_continuous(name ="Bitcoin Hourly Exchange Rate in USD",
                     breaks = c(0, 500, 1000, 1500, 2000, 2500, 3000)) +
  ggtitle("Bitcoin Daily Price",
          subtitle = "experienced the greatest surge in history this year") +
  labs(caption = "Data source: Coinbase API") +
  annotate(geom="text",
           x=as.POSIXct("2017-05-30"),
           y=675,
           label="2017-03-25 Surge starts",
           colour="#ff7575",
           fontface="bold",
           alpha=.85) +
  theme(panel.grid.major.y = element_line( size=.1, color="#666666"),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size=18,
                                  family = "Helvetica",
                                  colour = "#3a3a3a",
                                  face = "bold"),
        plot.subtitle = element_text(size=12, family = "mono", colour = "#666666"),
        axis.title.y.right = element_text(color="#85bb65"),
        axis.text.y.right = element_text(color="#85bb65"),
        axis.text.y = element_text(color="#325a8c"),
        axis.title.y = element_text(color="#325a8c"),
        axis.ticks.y = element_line( size=.25, color="#666666"),
        axis.text.x = element_text(color="#7f7f7f"),
        legend.position = "none",
        plot.caption = element_text(size=11, family = "mono", colour = "#666666"),
        plot.margin = unit(c(2,2,2,2), "cm"))

line
```

\pagebreak

### Chart two
```{r, echo=FALSE, include=FALSE, result="hide"}
nodes <- read_csv("nodes_by_countries.csv")
```

```{r, fig.width = 8, fig.height = 12}
hist <- ggplot(data=nodes, aes(reorder(factor(country), node), node)) +
  geom_col(width=0.9, position=position_dodge(width=5), fill = "#999999") +
  geom_text(aes(label=node), color="#666666", position=position_dodge(width=0.9), vjust=0.35, hjust=-.25) +
  ggtitle("Bitcoin Nodes Distribution by Countries",
          subtitle = "Reachable nodes as of Oct 15, 2017") +
  ylab("") +
  xlab("") +
  labs(caption = "Data source: BitNodes.21.co") +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size=18,
                                  family = "Helvetica",
                                  colour = "#3a3a3a",
                                  face = "bold"),
        plot.subtitle = element_text(size=12, family = "mono", colour = "#666666"),
        axis.text.y = element_text(color="#325a8c"),
        axis.title.y = element_text(color="#325a8c"),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none",
        plot.caption = element_text(size=11, family = "mono", colour = "#666666"),
        plot.margin = unit(c(2,2,2,2), "cm")) +
  scale_x_discrete(expand=c(0,0)) +
  coord_flip()

hist
```

\pagebreak

## Chart Three
```{r, echo=FALSE, include=FALSE, result="hide"}
#coinbase_lim$volume_usd, coinbase_lim$volume_btc, 
ts <- xts(coinbase_lim$volume_usd, as.Date(coinbase_lim$unix_timestamp, "%Y-%m-%d"))
ts_d = apply.daily(ts, FUN=sum)
#ts_m = apply.monthly(ts_d, FUN=sum)
df_d <- data.frame(date=index(ts_d), coredata(ts_d))
colnames(df_d) <- c("date", "volume_usd")

ts_btc <- xts(coinbase_lim$volume_btc, as.Date(coinbase_lim$unix_timestamp, "%Y-%m-%d"))
ts_d_btc = apply.daily(ts_btc, FUN=sum)
#ts_m_btc = apply.monthly(ts_d_btc, FUN=sum)
df_d_btc <- data.frame(date=index(ts_d_btc), coredata(ts_d_btc))
colnames(df_d_btc) <- c("date", "volume_btc")

df <- left_join(df_d, df_d_btc)

df_btc <- df[, c("date", "volume_btc")]
df_btc$unit <- "Bitcoin"
colnames(df_btc) <- c("date", "volume", "unit")

df_usd <- df[, c("date", "volume_usd")]
df_usd <- mutate(df_usd, volume_usd = volume_usd/1000)
df_usd$unit <- "Thousand USD"
colnames(df_usd) <- c("date", "volume", "unit")

agg <- rbind(df_btc, df_usd)
```


```{r, fig.width = 16, fig.height = 12}
# note: color #85bb65 is nicknamed as "dollar bill green".
#       color #FF9900 is the orange from bitcoin logo.
area <- ggplot(data=agg, aes(x = as.POSIXct(date), y = volume, fill = as.factor(unit))) +
  geom_area(position = "stack") +
  scale_x_datetime(date_breaks = "3 months", expand=c(0,0)) +
  scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 2000, 3000)) + 
  scale_fill_manual(values = alpha(c("#FF9900", "#85bb65")), 0.2) +
  ggtitle("Bitcoin Daily Trading Volume",
          subtitle = "in number of bitcoin & Thousands of USD") +
  labs(caption = "Data source: Coinbase API",
       x = "",
       y = "Volume") +
  theme(panel.grid.major.y =element_line( size=.1, color="#999999"),
        panel.grid.major.x = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(size=18,
                                  family = "Helvetica",
                                  colour = "#3a3a3a",
                                  face = "bold"),
        plot.subtitle = element_text(size=12, family = "mono", colour = "#666666"),
        axis.text.y = element_text(color="#666666"),
        axis.title.y = element_text(color="#666666"),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(size=10, family = "mono", colour = "#666666"),
        legend.title = element_blank(),
        legend.position = c(0.1, 0.9),
        plot.caption = element_text(size=11, family = "mono", colour = "#666666"),
        plot.margin = unit(c(2,2,2,2), "cm"))

area
```